### 2016-06-24 publications/Journal/1-introduction.tex
More fixes, resolve the unnumbered referencing of sections.
Signed-off-by: Ben Cardoen <bencardoen@gmail.com>

The remainder of this paper is organized as follows:
Section~\ref{sec:2-background} introduces the necessary background on synchronization protocols.
Section~\ref{sec:3-features} elaborates on our design that enables this flexibility.
In Section~\ref{sec:4-performance}, we evaluate performance of our tool by comparing its different synchronization protocols, and by comparing to adevs.
Related work is discussed in Section~\ref{sec:5-related-work}.
Section~\ref{sec:6-conclusion} concludes the paper and gives future work.

>>>>

%Usage of Hyperref with hardcoded label is hardly ideal, but ~\ref doesn't work if the layout disables numbered sections, so a hack is needed. The alternative is \nameref but that isn't all that readable.
The remainder of this paper is organized as follows:
Section \hyperref[sec:2-background]{2} introduces the necessary background on synchronization protocols.
Section~\hyperref[sec:3-features]{3} elaborates on our design that enables this flexibility.
In Section~\hyperref[sec:4-performance]{4}, we evaluate performance of our tool by comparing its different synchronization protocols, and by comparing to adevs.
Related work is discussed in Section~\hyperref[sec:5-related-work]{5}.
Section~\hyperref[sec:6-conclusion]{6} concludes the paper and gives future work.

### 2016-06-24 publications/Journal/3-features.tex
More fixes, resolve the unnumbered referencing of sections.
Signed-off-by: Ben Cardoen <bencardoen@gmail.com>

As we will demonstrate in Section~\ref{sec:4-subsec:overhead-pgraph:memory} memory allocation is a performance bottleneck in simulation with high event frequency, especially in parallel. Dxex has a modular allocation interface that separates the object creation in kernels and models from the actual implementation. This allows us by using a single compilation parameter to change the allocation strategy per kernel type. This interface allows instrumentation of memory usage. Not only useful to detect memory leaks should this be a concern, it can also be of use to a model author to trace event creation in large models. A model author can rely on the default implementation as the fastest common configuration verified by the benchmarks.

@@

A model author can specify which kernel a model should be allocated to, should such manual intervention be required. This is handled by the default model allocator. If no preference is given a simple striping scheme is followed. It should be obvious that this is not sufficient in most simulations. In \ref{sec:4-performance} we analyze allocation schemes and their performance impact. By overriding the default allocator a model author can tune the allocation scheme to maximize parallel speedup. This interface can be in future work also be used to employ graph algorithms for an automatic allocation scheme, for example one that avoids cycles in the resulting kernel dependency graph.

@@

\ref{PholdTreeallocation}. Dotted lines represent connections between models that did not see events exchanged, directed edges represents events with frequency as weight and color to distinguis inter versus intra kernel events.

@@

	\includegraphics[width=\plotfraction\columnwidth, height=8cm, keepaspectratio]{fig/pholdtreed1n3t5000.eps}

@@

	\includegraphics[width=\plotfraction\columnwidth,  height=8cm, keepaspectratio]{fig/pholdtreed1n3t5000c4BFS.eps}

@@

	\includegraphics[width=\plotfraction\columnwidth,  height=8cm, keepaspectratio]{fig/pholdtreed1n3t5000c4DFS.eps}

>>>>

As we will demonstrate in ~\nameref{sec:4-subsec:overhead-pgraph:memory} memory allocation is a performance bottleneck in simulation with high event frequency, especially in parallel. Dxex has a modular allocation interface that separates the object creation in kernels and models from the actual implementation. This allows us by using a single compilation parameter to change the allocation strategy per kernel type. This interface allows instrumentation of memory usage. Not only useful to detect memory leaks should this be a concern, it can also be of use to a model author to trace event creation in large models. A model author can rely on the default implementation as the fastest common configuration verified by the benchmarks.

@@

A model author can specify which kernel a model should be allocated to, should such manual intervention be required. This is handled by the default model allocator. If no preference is given a simple striping scheme is followed. It should be obvious that this is not sufficient in most simulations. In Section \nameref{sec:4-performance} we analyze allocation schemes and their performance impact. By overriding the default allocator a model author can tune the allocation scheme to maximize parallel speedup. This interface can be in future work also be used to employ graph algorithms for an automatic allocation scheme, for example one that avoids cycles in the resulting kernel dependency graph.

@@

\nameref{PholdTreeallocation} . Dotted lines represent connections between models that did not see events exchanged, directed edges represents events with frequency as weight and color to distinguish inter versus intra kernel events.

@@

	\includegraphics[width=\plotfraction\columnwidth, height=6cm, keepaspectratio]{fig/pholdtreed1n3t5000.eps}

@@

	\includegraphics[width=\plotfraction\columnwidth,  height=6cm, keepaspectratio]{fig/pholdtreed1n3t5000c4BFS.eps}

@@

	\includegraphics[width=\plotfraction\columnwidth,  height=6cm, keepaspectratio]{fig/pholdtreed1n3t5000c4DFS.eps}

### 2016-06-24 publications/Journal/parallel-performance.tex
More fixes, resolve the unnumbered referencing of sections.
Signed-off-by: Ben Cardoen <bencardoen@gmail.com>

% The real life
\subsubsection{PholdTree}
%move to pholdtree
We further verify that our contribution fulfills our projected use case: a single model that can be tweaked to favor either conservative or optimistic synchronization. We will demonstrate that allocation is critical to achieve a parallel speedup and that the parameters of the model will give either conservative or optimistic an advantage.
\paragraph*{Allocation}\label{PholdTreeallocation}
The PholdTree benchmark can be configured to use 2 allocation schemes: breadth first and depth first. 
The breadth first allocation scheme will result in kernels that will form a dependency chain with multiple branches, much like in the Queue model. 
Such a linear dependency chain can result in a parallel speedup as we demonstrated with the Queue model, but this is not always true as we will demonstrate in this section. 
A single kernel that has an unbalanced number of atomic models or models with unequal computation load in transition functions will slow down the remainder of the chain. This effect is also apparent if the OS unfairly schedules the thread a kernel runs on. In conservative this will lead to excessive polling of the other kernels' eot values, whereas in optimistic this will lead to a cascade of reverts since dependent kernels will simulate ahead of the slower kernel. In Figures \ref{fig:pholdtree_visualize_parBFS} and \ref{fig:pholdtree_visualize_parDFS} the simulation trace is visualized for both allocation schemes highlighting the remarks made in this paragraph.

	\includegraphics[width=\modelfraction\columnwidth]{fig/pholdtreeBFS.eps}
	\caption{PholdTree model breadth first allocation with 3 kernels.}
	\label{fig:PholdTree_model_bfs}

	\includegraphics[width=\modelfraction\columnwidth]{fig/pholdtreeDFS.eps}
	\caption{PholdTree model depth first allocation with 3 kernels.}
	\label{fig:PholdTree_model_dfs}
\end{figure}

@@

\paragraph*{Weak Scaling}
%Intro
In this paragraph we investigate dxex's parallel performance when the the number of atomic models distributed over a fixed number of kernels varies. From \ref{eq:pholdtreemodelcount} we know that increasing d will lead to a very rapid increase in models, we want to observe what happens in a closer to linear increase in model count per kernel, so we vary n. \\

% Explain why adevs is not plotted (its x100 slower)

Adevs' conservative implementation cannot handle the uncertainty (lack of non $\epsilon$ lookahead) here and is omitted from the comparison. Its parallel performance is two orders of magnitude slower than the sequential implementation. Upon investigation we conclude that this slowdown is caused by not implementing the \{begin/end\}lookahead() functions, which when not overridden in a model trigger an exception on each invocation. \\
This exception handling completely stalls the kernel. We do not implement this function in our PholdTree variant for adevs since this would force us to implement state saving in the model code and not use the provided state saving functionality in the kernel, as is done in dxex or PythonPDEVS's optimistic.\\
In dxex state saving is never required for a conservative simulation, this is handled transparently for the user who need not implement this behaviour. 
By using the state saving technique inside the model code we feel we would no longer compare identical models across different synchronization algorithms, although we do not doubt adevs' increased performance when these functions are overridden.
\\

% Optimistic
In Figure \ref{fig:PholdTree_plot_weaknopt} we observe that dxex's optimistic kernels are slightly sensitive to the number of atomic models allocated to them. In the worst case with 2 kernels there is no real speedup observable, as soon as the number of kernels increases we see a converging speedup trend, that except for 32 kernels is almost constant despite the increase in n. Note that only depth first allocated kernels are measured here, breadth first as we can see in \ref{fig:PholdTree_plot_alloc_high} has no speedup advantage. The probability parameter is kept at 0.1 for the same reason, it only induces a linear increase in load.
\\

>>>>

% The real life
\subsubsection{PholdTree}
%move to pholdtree
We further verify that our contribution fulfills our projected use case: a single model that can be tweaked to favor either conservative or optimistic synchronization. We will demonstrate that allocation is critical to achieve a parallel speedup and that the parameters of the model will give either conservative or optimistic an advantage.
\paragraph*{Allocation}\label{PholdTreeallocation}
The PholdTree benchmark can be configured to use 2 allocation schemes: breadth first and depth first. 
The breadth first allocation scheme will result in kernels that will form a dependency chain with multiple branches, much like in the Queue model. 
Such a linear dependency chain can result in a parallel speedup as we demonstrated with the Queue model, but this is not always true as we will demonstrate in this section. 
A single kernel that has an unbalanced number of atomic models or models with unequal computation load in transition functions will slow down the remainder of the chain. This effect is also apparent if the OS unfairly schedules the thread a kernel runs on. In conservative this will lead to excessive polling of the other kernels' eot values, whereas in optimistic this will lead to a cascade of reverts since dependent kernels will simulate ahead of the slower kernel. In Figures \ref{fig:pholdtree_visualize_parBFS} and \ref{fig:pholdtree_visualize_parDFS} the simulation trace is visualized for both allocation schemes highlighting the remarks made in this paragraph.
%
%\begin{figure}
%	\center
%	
%	\includegraphics[width=\modelfraction\columnwidth]{fig/pholdtreeBFS.eps}
%	\caption{PholdTree model breadth first allocation with 3 kernels.}
%	\label{fig:PholdTree_model_bfs}
%	
%	\includegraphics[width=\modelfraction\columnwidth]{fig/pholdtreeDFS.eps}
%	\caption{PholdTree model depth first allocation with 3 kernels.}
%	\label{fig:PholdTree_model_dfs}
%\end{figure}

@@

\paragraph*{Weak Scaling}
%Intro
In this paragraph we investigate dxex's parallel performance when the the number of atomic models distributed over a fixed number of kernels varies. From \ref{eq:pholdtreemodelcount} we know that increasing d will lead to a very rapid increase in models, we want to observe what happens in a closer to linear increase in model count per kernel, so we vary n.

% Explain why adevs is not plotted (its x100 slower)
Adevs' conservative implementation cannot handle the uncertainty (lack of non $\epsilon$ lookahead) here and is omitted from the comparison. Its parallel performance is two orders of magnitude slower than the sequential implementation. Upon investigation we conclude that this slowdown is caused by not implementing the \{begin/end\}lookahead() functions, which when not overridden in a model trigger an exception on each invocation. 
This exception handling completely stalls the kernel. We do not implement this function in our PholdTree variant for adevs since this would force us to implement state saving in the model code and not use the provided state saving functionality in the kernel, as is done in dxex or PythonPDEVS's optimistic.
In dxex state saving is never required for a conservative simulation, this is handled transparently for the user who need not implement this behaviour. 
By using the state saving technique inside the model code we feel we would no longer compare identical models across different synchronization algorithms, although we do not doubt adevs' increased performance when these functions are overridden.

% Optimistic
In Figure \ref{fig:PholdTree_plot_weaknopt} we observe that dxex's optimistic kernels are slightly sensitive to the number of atomic models allocated to them. In the worst case with 2 kernels there is no real speedup observable, as soon as the number of kernels increases we see a converging speedup trend, that except for 32 kernels is almost constant despite the increase in n. Note that only depth first allocated kernels are measured here, breadth first as we can see in \ref{fig:PholdTree_plot_alloc_high} has no speedup advantage. The probability parameter is kept at 0.1 for the same reason, it only induces a linear increase in load.
